Combinatorics of the geometry of Wilson loop diagrams II: Grassmann necklaces, dimensions, and denominators

### 2020-12-30 GWLDCII.tex
maybe final edits?

Meanwhile, a different body of work has studied SYM $N=4$ theory from the point of view of Feynman integrals in twistor space \cite{Adamo:2011pv,Adamo:2012xe,  Boels:2007qn, Bullimore:2010pj}. These integrals are calculated via holomorphic Wilson loops. As with the Amplituhedron, each Wilson loop represents the sum of a family of $N^kMHV$ Feynman diagrams. Each of the Feynman integrands in this sum associates a volume form to a convex space, parallell to the Amplituhedron \cite{CachazoSvrcekWitten, Hodges2013}. However, the geometric objects decomposing the Wilson loops are different from the geometric object decomposing the Amplituhedron.

@@

With these definitions in hand, we can impose some conditions on the density and behavior of the propagators of Wilson loop diagrams.

@@

All three of these conditions arise from the physical interpretation of Wilson loop diagrams.  The first condition comes from the fact that these diagrams represent $N^k$MHV interctions, where $k +4 <n$, see for instance the explanation in the introduction of \cite{Adamo:2011pr}.  The second condition can be derived from the integrals associated to the diagrams, as in \cite[Theorem 1.13]{wilsonloop}. Note that the second condition prohibits propagators that connect two adjacent edges. In our definition, $\cP$ is a set so we may not have two propagators starting and ending on the same pair of edges, but we have lost nothing of physical value by not allowing multiple propagators ending on the same pair of edges since these would also fail the second condition. The third condition comes from the fact that these Wilson loop diagrams are a representation of planar Feynman diagrams of $N=4$ SYM theory. For a description of the relationship between planar Feynman diagram and planar Wilson loops, see the discussion in \cite[section 3]{LipsteinMason}.

@@

This would certainly hold if the singularities in the integrands $\Omega(W)$ cancel out in the sum.  The singularities coming from the zeros of $R(W)$ are the spurious poles of the theory and the physical theory tells us that they should cancel. 
%(the spurious poles of the theory) cancel out in the sum; in particular, if all poles arising from $R(W)$ must cancel.
Specifically, it is conjectured that these poles cancel on the boundaries of the positroid cells $\Sigma(W)$, analogous to the analysis in \cite{Arkani-Hamed:2013jha}. This has been checked for particular classes of examples (e.g. \cite{casestudy, Amplituhedronsquared, HeslopStewart}) but has not been proven in general.

@@

A reader interested in studying these integrals further is invited to read \cite{casestudy, HeslopStewart, Amplituhedronsquared} in order to understand the connection to the scattering amplitude of $N=4$ SYM theory, the cancelation of spurious poles of specific instances of these integrals as well as explicit amplitude calculations.

@@

  Because the nonzero entries of $C(W)$ are independent indeterminants, $J$ is an independent set if and only if there is some choice of $|J|$ nonzero entries of $C(W)$ one in each row associated to an element of $J$ and each in different columns. Each entry in $C(W)$ identifies a propagator by the row of the entry and a vertex by the column of the entry.  The entry is nonzero if and only if the propagator is supported on that vertex.  Consequently, take a choice of $|J|$ nonzero entries of $C(W)$, one in each row associated to an element of $J$ and each in different columns.  Such a choice is equivalent to an assignment of the propagators of $J$ to supporting vertices so that no two are assigned to the same vertex.  Such an assignment of the propagators of $J$ to supporting vertices is exactly a map $f$ as described in the statement, hence proving the result.
\end{proof}

@@

By relating the behavior of the propagators in $W$ to the matroidal properties of $M(W)$, Agarwala and Marin-Amat also showed that $M(W)$ is a positroid whenever $W$ is admissible \cite[Corollary 3.39]{wilsonloop}.

@@

Before we can describe the algorithm for extracting the Grassmann necklace of $M(W)$ from the Wilson loop diagram $W$, we require some initial results about the behavior of propagators in admissible Wilson loop diagrams.

@@

In this section, we give an algorithm for passing directly from a Wilson loop diagram to its Grassmann necklace. This not only greatly simplifies the previously known process to obtain a positroid cell from a Wilson loop diagram, but also allows us to relate the behavior of the positroid $M(W)$ directly to the configuration of propagators in $W$. In Section~\ref{sec poles} we use these results to show that the denominator $R(W)$ of the integrand associated to a Wilson loop diagram can be written in terms of the Grassmann necklace of said diagram.

@@

\begin{rmk}\label{rmk algorithm locally same}
The following observation about the local behavior of the algorithm will be very useful in many of the arguments below.

Suppose we are going through the algorithm to construct $I_i^W$ for some $i$, and we are currently at step $j$ of the loop. Consider the diagram $W'$ formed by removing from $W$ all propagators that have already contributed to $I_i^W$. (This replicates the behavior of the algorithm, which removes each propagator once it contributes to $I_i^W$.)  Now suppose for some other diagram $V$ on the same vertex set we are constructing $I_m^V$ for some $m$, and we are at step $j$ of the loop, for the same $j$ as above.  Again obtain $V'$ from $V$ by removing all propagators assigned so far.

@@

When the firt possibility occurs, observe that we must have another propagator $q = (c,b)$ with $c<_i b$ and $I_i(q) = b+1$, because otherwise $I_i$ would assign $p$ to $b+1$. Now $q$ also contributes the $<_i$-maximal vertex in its support to $I_i$ and is an instance of the second possibility. Therefore, it suffices to show that no admissible Wilson loop diagram admits the configuration illustrated in Case 2 of Figure~\ref{fig:no fourth vertex}.

@@

Marcott in \cite[Theorem 8.4]{WLDdim} also gave a proof of the $3|\cP|$-dimensionality of $M(W)$ which is much more geometric. However, it is not easy to track the effect of a particular propagator in that proof. By contrast, our approach is much more constructive, and elucidates how each additional propagator contributes to tthe dimension. By combining Algorithms~\ref{alg:put GN on WLD} and \ref{alg:GN to Le} (converting from Wilson loop diagram to Grassmann necklace, and Grassmann necklace to Le diagram respectively) we explicitly describe the effect of adding another propagator to a Wilson loop diagram in terms of the plusses of the associated Le diagrams. Recall that the dimension of a positroid cell is equal to the number of plusses in its associated Le diagram \cite[Theorem 6.5]{Postnikov}.  Thus understanding the effect of adding another propagator is sufficient to give a recursive proof of the $3|\cP|$-dimensionality of the cells.

@@

Each $r_i$ is the determinant of a minor determined by certain propagators of $W$. In particular, if a propagator $p$ contributes one value to the $i^{th}$ element of the Grassmann necklace and a different value to the $(i-1)^{th}$ element, then $p$ is one of the propagators contibuting to the minor. Specifically, the row associated to $p$ and the column associated to the value that $p$ contributes to the $i^{th}$ Grassmann Necklace element appear in the minor.
%That is, each $r_i$ is the determinant of the minor determined by the propagators (rows) that contribute different values (columns) to the $i^{th}$ element of the Grassmann necklace than they did to the $(i-1)^{th}$. 
Since each $r_i$ is simply the determinant of a particular submatrix of $C(W)$ and an ordering on the propagators has been fixed, the sign of each $r_i$ is well defined.

>>>>

Meanwhile, a different body of work has studied SYM $N=4$ theory from the point of view of Feynman integrals in twistor space \cite{Adamo:2011pv,Adamo:2012xe,  Boels:2007qn, Bullimore:2010pj}. These integrals are calculated via holomorphic Wilson loops. As with the Amplituhedron, each Wilson loop represents the sum of a family of $N^kMHV$ Feynman diagrams. Each of the Feynman integrands in this sum associates a volume form to a convex space, parallel to the Amplituhedron \cite{CachazoSvrcekWitten, Hodges2013}. However, the geometric objects decomposing the Wilson loops are different from the geometric object decomposing the Amplituhedron.

@@

With these definitions in hand, we can impose some conditions on the density and behaviour of the propagators of Wilson loop diagrams.

@@

All three of these conditions arise from the physical interpretation of Wilson loop diagrams.  The first condition comes from the fact that these diagrams represent $N^k$MHV interactions, where $k +4 <n$, see for instance the explanation in the introduction of \cite{Adamo:2011pr}.  The second condition can be derived from the integrals associated to the diagrams, as in \cite[Theorem 1.13]{wilsonloop}. Note that the second condition prohibits propagators that connect two adjacent edges. In our definition, $\cP$ is a set so we may not have two propagators starting and ending on the same pair of edges, but we have lost nothing of physical value by not allowing multiple propagators ending on the same pair of edges since these would also fail the second condition. The third condition comes from the fact that these Wilson loop diagrams are a representation of planar Feynman diagrams of $N=4$ SYM theory. For a description of the relationship between planar Feynman diagram and planar Wilson loops, see the discussion in \cite[section 3]{LipsteinMason}.

@@

This would certainly hold if the singularities in the integrands $\Omega(W)$ cancel out in the sum.  The singularities coming from the zeros of $R(W)$ are the spurious poles of the theory, as mentioned in the introduction. The physical theory tells us that the spurious poles should cancel. 
%(the spurious poles of the theory) cancel out in the sum; in particular, if all poles arising from $R(W)$ must cancel.
Specifically, it is conjectured that these poles cancel on the boundaries of the positroid cells $\Sigma(W)$, analogous to the analysis in \cite{Arkani-Hamed:2013jha}. This has been checked for particular classes of examples (e.g. \cite{casestudy, Amplituhedronsquared, HeslopStewart}) but has not been proven in general.

@@

A reader interested in studying these integrals further is invited to read \cite{casestudy, HeslopStewart, Amplituhedronsquared} in order to understand the connection to the scattering amplitude of $N=4$ SYM theory, the cancellation of spurious poles of specific instances of these integrals as well as explicit amplitude calculations.

@@

  Because the nonzero entries of $C(W)$ are independent indeterminates, $J$ is an independent set if and only if there is some choice of $|J|$ nonzero entries of $C(W)$ one in each row associated to an element of $J$ and each in different columns. Each entry in $C(W)$ identifies a propagator by the row of the entry and a vertex by the column of the entry.  The entry is nonzero if and only if the propagator is supported on that vertex.  Consequently, take a choice of $|J|$ nonzero entries of $C(W)$, one in each row associated to an element of $J$ and each in different columns.  Such a choice is equivalent to an assignment of the propagators of $J$ to supporting vertices so that no two are assigned to the same vertex.  Such an assignment of the propagators of $J$ to supporting vertices is exactly a map $f$ as described in the statement, hence proving the result.
\end{proof}

@@

By relating the behaviour of the propagators in $W$ to the matroidal properties of $M(W)$, Agarwala and Marin-Amat also showed that $M(W)$ is a positroid whenever $W$ is admissible \cite[Corollary 3.39]{wilsonloop}.

@@

Before we can describe the algorithm for extracting the Grassmann necklace of $M(W)$ from the Wilson loop diagram $W$, we require some initial results about the behaviour of propagators in admissible Wilson loop diagrams.

@@

In this section, we give an algorithm for passing directly from a Wilson loop diagram to its Grassmann necklace. This not only greatly simplifies the previously known process to obtain a positroid cell from a Wilson loop diagram, but also allows us to relate the behaviour of the positroid $M(W)$ directly to the configuration of propagators in $W$. In Section~\ref{sec poles} we use these results to show that the denominator $R(W)$ of the integrand associated to a Wilson loop diagram can be written in terms of the Grassmann necklace of said diagram.

@@

\begin{rmk}\label{rmk algorithm locally same}
The following observation about the local behaviour of the algorithm will be very useful in many of the arguments below.

Suppose we are going through the algorithm to construct $I_i^W$ for some $i$, and we are currently at step $j$ of the loop. Consider the diagram $W'$ formed by removing from $W$ all propagators that have already contributed to $I_i^W$. (This replicates the behaviour of the algorithm, which removes each propagator once it contributes to $I_i^W$.)  Now suppose for some other diagram $V$ on the same vertex set we are constructing $I_m^V$ for some $m$, and we are at step $j$ of the loop, for the same $j$ as above.  Again obtain $V'$ from $V$ by removing all propagators assigned so far.

@@

When the first possibility occurs, observe that we must have another propagator $q = (c,b)$ with $c<_i b$ and $I_i(q) = b+1$, because otherwise $I_i$ would assign $p$ to $b+1$. Now $q$ also contributes the $<_i$-maximal vertex in its support to $I_i$ and is an instance of the second possibility. Therefore, it suffices to show that no admissible Wilson loop diagram admits the configuration illustrated in Case 2 of Figure~\ref{fig:no fourth vertex}.

@@

Marcott in \cite[Theorem 8.4]{WLDdim} also gave a proof of the $3|\cP|$-dimensionality of $M(W)$ which is much more geometric. However, it is not easy to track the effect of a particular propagator in that proof. By contrast, our approach is much more constructive, and elucidates how each additional propagator contributes to the dimension. By combining Algorithms~\ref{alg:put GN on WLD} and \ref{alg:GN to Le} (converting from Wilson loop diagram to Grassmann necklace, and Grassmann necklace to Le diagram respectively) we explicitly describe the effect of adding another propagator to a Wilson loop diagram in terms of the plusses of the associated Le diagrams. Recall that the dimension of a positroid cell is equal to the number of plusses in its associated Le diagram \cite[Theorem 6.5]{Postnikov}.  Thus understanding the effect of adding another propagator is sufficient to give a recursive proof of the $3|\cP|$-dimensionality of the cells.

@@

Each $r_i$ is the determinant of a minor determined by certain propagators of $W$. In particular, if a propagator $p$ contributes one value to the $i^{th}$ element of the Grassmann necklace and a different value to the $(i-1)^{th}$ element, then $p$ is one of the propagators contributing to the minor. Specifically, the row associated to $p$ and the column associated to the value that $p$ contributes to the $i^{th}$ Grassmann Necklace element appear in the minor.
%That is, each $r_i$ is the determinant of the minor determined by the propagators (rows) that contribute different values (columns) to the $i^{th}$ element of the Grassmann necklace than they did to the $(i-1)^{th}$. 
Since each $r_i$ is simply the determinant of a particular submatrix of $C(W)$ and an ordering on the propagators has been fixed, the sign of each $r_i$ is well defined.
